Field-informed Reinforcement Learning for Learning Large-Scale Collective Tasks

### 2023-05-16 paper.tex
chore: more on notation

On these minimal operators, 
 it is then possible to build \emph{self-organizing} coordination blocks.
 One of the founding blocks is the \emph{gradient} essential for 
 \emph{information} flows~\cite{DBLP:conf/saso/WolfH07}.
%
This operator generates a numerical field that represents the minimum distance from a source zone. 
%
In other words, it maps a Boolean field (where ``true'' indicates the presence of a source node and ``false'' indicates its absence) to a distance field that indicates the proximity to the nearest source. In ScaFi (a scala implementation of aggregate computing) it is simply expressed as:
\begin{lstlisting}
def gradient(source: Boolean): Double
\end{lstlisting}
Along the gradient, it is possible to cast information within the system, accumulating value during the programming. This pattern is called \emph{gradient-cast} (or \emph{G}) 
\begin{lstlisting}
def G[V](source: Boolean, value: V, acc: V => V): V
\end{lstlisting}
\ga{more description here}
this is used in the use case to accumulate the distance from the phenomena that is tracked
\lukas{still not sure how the coordination works or if it has been explained in the above paragraph.}
\subsection{Graph Neural Networks}
%Over the years, several different neural networks have been proposed to solve specific tasks, ranging from 
%  simple \ac{MLP}~\cite{werbos1975beyond} to \ac{CNN}~\cite{oshea2015introduction}.
%  \lukas{more examples? }
%%
%Whereas \ac{CNN} is designed to be used with image and spatial-like data (e.g., audio), 
 \ac{GNN} is a novel neural network model used to process graph-structured data with deep learning approaches.
%
Let $G = (V, E)$ a graph  
where $E$ defines the neighbourhood relations for each participating node, 
and $V$ identifies the nodes present in the graph. %at that time.  \lukas{no time in the definition}
% 
Each node $v \in V$ is associated with an observation (or feature set) $f_v$.
%\lukas{Over time this graph $G$ can change and therefore we define $G_t$ as the graph $G$ at a specific time $t$.} 
%
Given $f_v$, %a feature vector associated with each node $v \in V$, 
 the goal of a \ac{GNN} is to learn the node embedding $h_v$ for each node $v \in V$. The node embedding $h_v$ describes the node in the network and summarises the geometric properties of the graph in this location, allowing for comparison of various nodes in the graph.
 In modern \acp{GNN}, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $\mathcal{N}_G(v)$,
  and then combining it with the node's current embedding $h_v$.
 The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. With each traversal of an edge $e \in E$ from $v$, the number of layers increased
Formally, a \ac{GNN} can be defined by two phases:
\begin{equation}
a_{v}^{(k)}=\operatorname{AGGREGATE}^{(k)}\left(\left\{h_{u}^{(k-1)}: u \in \mathcal{N}_G(v)\right\}\right)   
\end{equation}

\begin{equation}
h_{v}^{(k)}=\operatorname{COMBINE}^{(k)}\left(h_{v}^{(k-1)}, a_{v}^{(k)}\right)
\end{equation}

where $h_{v}^{k}$ is the embedding of node $v$ within the $k$-th layer, 
 $\mathcal{N}_G(v)$ is the set of neighbours of node $v$ computed from $E$.
%
$h_v{v}^0$ is the initial embedding of node $v$ and it is usually set to the node's feature vector $f_v$.
%
The differential part comes into play in the $\operatorname{COMBINE}$ function, 
 which is usually a differentiable function such as a neural network layer.
%
$\operatorname{AGGREGATE} $ instead is a function that aggregates the information from the neighbours of a node $v$
  and it could be a simple sum, max or sum of products. More complex aggregation functions are available~\cite{pellegrini2020learning}.
%
This formulation allows GNNs to effectively process and extract features from graph-structured data by iteratively aggregating and transforming information from the node's neighbours.
\lukas{do we need definitions of $o_v$, $h_v^*$ over time $t$?}

>>>>

On these minimal operators, 
 it is then possible to build \emph{self-organizing} coordination blocks.
 One of the founding blocks is the \emph{gradient} essential for 
 \emph{information} flows~\cite{DBLP:conf/saso/WolfH07}.
%
This operator generates a numerical field that represents the minimum distance from a source zone. 
%
In other words, it maps a Boolean field (where ``true'' indicates the presence of a source node and ``false'' indicates its absence) to a distance field that indicates the proximity to the nearest source. In ScaFi (a scala implementation of aggregate computing) it is simply expressed as:
\begin{lstlisting}
def gradient(source: Boolean): Double
\end{lstlisting}
Along the gradient, it is possible to cast information within the system, accumulating value during the programming. This pattern is called \emph{gradient-cast} (or \emph{G}) 
\begin{lstlisting}
def G[V](source: Boolean, value: V, acc: V => V): V
\end{lstlisting}
\ga{more description here}
this is used in the use case to accumulate the distance from the phenomena that is tracked
\lukas{still not sure how the coordination works or if it has been explained in the above paragraph.}
\subsection{Graph Neural Networks}
%Over the years, several different neural networks have been proposed to solve specific tasks, ranging from 
%  simple \ac{MLP}~\cite{werbos1975beyond} to \ac{CNN}~\cite{oshea2015introduction}.
%  \lukas{more examples? }
%%
%Whereas \ac{CNN} is designed to be used with image and spatial-like data (e.g., audio), 
 \ac{GNN} is a novel neural network model used to process graph-structured data with deep learning approaches.
%
Let $G = (V, E)$ a graph  
where $E$ defines the neighbourhood relations for each participating node, 
and $V$ identifies the nodes present in the graph. %at that time.  \lukas{no time in the definition}
% 
Each node $v \in V$ is associated with an observation (or feature set) $f_v$.
For the sake of simplicity, 
 we thereafter describe $G_f$ as a graph that contains the feature set $f_v$ for each node $v \in V$.
%\lukas{Over time this graph $G$ can change and therefore we define $G_t$ as the graph $G$ at a specific time $t$.} 
%
Given $f_v$, %a feature vector associated with each node $v \in V$, 
 the goal of a \ac{GNN} is to learn the node embedding $h_v$ for each node $v \in V$. The node embedding $h_v$ describes the node in the network and summarises the geometric properties of the graph in this location, allowing for comparison of various nodes in the graph.
 In modern \acp{GNN}, the node embedding $h_v$ is computed by aggregating information from the node's neighbours $\mathcal{N}_G(v)$,
  and then combining it with the node's current embedding $h_v$.
 The graph is further partitioned into layers with the node $v$ at its bottom layer $0$. With each traversal of an edge $e \in E$ from $v$, the number of layers increased
Formally, a \ac{GNN} can be defined by two phases:
\begin{equation}
a_{v}^{(k)}=\operatorname{AGGREGATE}^{(k)}\left(\left\{h_{u}^{(k-1)}: u \in \mathcal{N}_G(v)\right\}\right)   
\end{equation}

\begin{equation}
h_{v}^{(k)}=\operatorname{COMBINE}^{(k)}\left(h_{v}^{(k-1)}, a_{v}^{(k)}\right)
\end{equation}
where $h_{v}^{k}$ is the embedding of node $v$ within the $k$-th layer, 
 $\mathcal{N}_G(v)$ is the set of neighbours of node $v$ computed from $E$.
%
$h_v{v}^0$ is the initial embedding of node $v$ and it is usually set to the node's feature vector $f_v$.
%
The differential part comes into play in the $\operatorname{COMBINE}$ function, 
 which is usually a differentiable function such as a neural network layer.
%
$\operatorname{AGGREGATE} $ instead is a function that aggregates the information from the neighbours of a node $v$
  and it could be a simple sum, max or sum of products. More complex aggregation functions are available~\cite{pellegrini2020learning}.
%
Thereafter, we express the application of a \ac{GNN} to a graph $G_f$ as:
\begin{equation}
    GNN(G_f) = \{h_v^{(k)}: v \in V, k \in \mathbb{N}\}
\end{equation} 
This formulation allows GNNs to effectively process and extract features from graph-structured data by iteratively aggregating and transforming information from the node's neighbours.
\lukas{do we need definitions of $o_v$, $h_v^*$ over time $t$?}
